---
title: "Example Code"
author: "Michelle Theodory"
date: "1/7/2022"
output: html_document
---
Loading in appropriate packages:
```{r, results=FALSE, warning=FALSE, message=FALSE}
library(psych)
library(gamlss)
library(tidyverse)
library(knitr)
library(here)
library(mgcv)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(xtable)
```

Setting path for saving figures: 
```{r}
figure_path <- "~/Desktop"
```

Loading in example data:
```{r, echo=TRUE, results=FALSE, warning=FALSE, message=FALSE}
personsum <- read_csv('~theodorymm/downloads/personsummary.csv')
daysum <- read_csv('~theodorymm/downloads/daysummary.csv')
pubertal_m <- read_csv('/Users/theodorymm/downloads/malepuberty.csv', na = c('NULL'))
pubertal_f <- read_csv('/Users/theodorymm/downloads/femalepuberty.csv', na = c('NULL'))
impairment <- read_csv('/Users/theodorymm/Downloads/CIS_score.csv')
demo <- read_csv('/Users/theodorymm/downloads/demos.csv')
ethnicity <- read_csv('/Users/theodorymm/Downloads/moredemos.csv')
BMI <- read_csv('/Users/theodorymm/downloads/cdc_data.csv')
diagnosis <- read_csv('/Users/theodorymm/downloads/Diagnoses.csv')
```

Renaming column names and subsetting data:

```{r}
impairment <- impairment %>% rename(ID = EID)

diagnosis <- diagnosis %>% rename(ID = EID)
diagnosis <- diagnosis[-c(1:7, 13, 25:26, 29, 44, 90:158)]

pubertal <- bind_rows(pubertal_m, pubertal_f)
pubertal <- pubertal %>% rename(ID = EID)

ethnicity <- subset(ethnicity, select=c(`PreInt_Demos_Fam,Child_CountryOrigin`, `PreInt_Demos_Fam,Child_Ethnicity`, `PreInt_Demos_Fam,Child_Race`, `PreInt_Demos_Fam,Child_Race_Other`, `PreInt_Demos_Fam,EID`))
names(ethnicity)[5] <- "ID"

names(BMI)[33] <- "ID"
BMI <- subset(BMI, select=c(ID, weight, height, bmi, wapct, waz, haz, hapct, bmipct, bmiz))

demo <- demo %>% rename(ID = EID)
demo <- demo%>% dplyr::select(ID, Age, Sex)

plotdata <- personsum %>% dplyr::select(ID, ACC_day_mg_pla, dur_day_total_IN_min_pla, dur_day_total_LIG_min_pla,
                                        dur_day_total_MOD_min_pla, dur_day_total_VIG_min_pla,  dur_spt_sleep_min_pla, sleep_efficiency_pla, L5TIME_num_wei, M10TIME_num_wei, M5TIME_num_wei, L10TIME_num_wei)

plotdata <- plotdata %>% mutate(dur_day_total_MVPA = dur_day_total_MOD_min_pla + dur_day_total_VIG_min_pla)
```

Merging the data, renaming/scoring and creating appropriate variables, and some data cleaning:

```{r}
plotdata <- merge(plotdata, pubertal, by="ID", all.x=TRUE) 
plotdata <- merge(plotdata, demo, by="ID", all.x=TRUE)
plotdata <- merge(plotdata, BMI, by= "ID", all.x=TRUE)
plotdata <- left_join(plotdata, ethnicity, by = "ID", all=T)
plotdata <- left_join(plotdata, impairment, by="ID", all=T)
plotdata <- merge(plotdata, diagnosis, by="ID", all.x=TRUE)

plotdata <- plotdata %>%
  mutate(PPS_M_Score = ifelse(PPS_M_Score == 3, 1, 
                              ifelse(PPS_M_Score %in% c(4,5), 2,
                                     ifelse(PPS_M_Score %in% c(6,7,8), 3,
                                            ifelse(PPS_M_Score %in% c(9,10, 11), 4,
                                                   ifelse(PPS_M_Score == 12, 5, NA))))),
         PPS_Score_Combined = ifelse(is.na(PPS_M_Score), PPS_F_Score, PPS_M_Score))

missing_demo <- plotdata %>% filter(is.na(Sex))
plotdata <- filter(plotdata, ! ID %in% missing_demo$ID)

plotdata$Sex <- as.factor(plotdata$Sex)
levels(plotdata$Sex) <- c("Males", "Females")

plotdata$PPS_F_06 <- as.factor(plotdata$PPS_F_06)
plotdata$PPS_M_04 <- as.factor(plotdata$PPS_M_04)
levels(plotdata$PPS_M_04) <- c("2","3","4","5")

plotdata <- 
  plotdata %>% 
  mutate(PPS_M_fac = as.character(PPS_M_04, levels=c("2","3","4","5"), labels= c("1 = not yet started changing", "2= has barely started changing", "3= voice change is definitely underway", "4= voice change seems completed")), 
         PPS_F_fac = as.character(PPS_F_06, levels=c("0","1"), labels=c("Without Menarche", "With Menarche")), 
         pubertal_combined = ifelse(Sex == "Males", PPS_M_fac, PPS_F_fac))

levels(plotdata$pubertal_combined) <- c("Without Menarche", "With Menarche", "No Voice Change", "Voice Change Barely Begun", "Voice Change Definitely Underway", "Voice Change Seems Completed")
```

Here, I am creating a combined race/ethnicity variable: 

```{r}
plotdata$hisp <- as.numeric(plotdata$`PreInt_Demos_Fam,Child_Ethnicity`==1)
plotdata$white <- as.numeric(plotdata$`PreInt_Demos_Fam,Child_Ethnicity`!=1 & plotdata$`PreInt_Demos_Fam,Child_Race`==0)
plotdata$black <- as.numeric(plotdata$`PreInt_Demos_Fam,Child_Ethnicity`!=1 & plotdata$`PreInt_Demos_Fam,Child_Race`==1)
plotdata$other <- as.numeric(plotdata$`PreInt_Demos_Fam,Child_Ethnicity`!= 1 & c(plotdata$`PreInt_Demos_Fam,Child_Race`==3 | plotdata$`PreInt_Demos_Fam,Child_Race`==4 | plotdata$`PreInt_Demos_Fam,Child_Race`== 5 | plotdata$`PreInt_Demos_Fam,Child_Race`==6 | plotdata$`PreInt_Demos_Fam,Child_Race`==7 | plotdata$`PreInt_Demos_Fam,Child_Race`== 8 | plotdata$`PreInt_Demos_Fam,Child_Race`==9 |plotdata$`PreInt_Demos_Fam,Child_Race`== 10 | plotdata$`PreInt_Demos_Fam,Child_Race`== 11)) 
plotdata$hi <- paste(plotdata$hisp, plotdata$white, plotdata$black, plotdata$other, sep = ",")
plotdata$hi <- as.factor(plotdata$hi)
plotdata$ethnic <- droplevels(x=plotdata$hi, exclude =  c("NA,0,NA,0", "0,NA,NA,NA", "NA,NA,0,0", "NA,NA,NA,NA", "NA,0,0,NA"))
levels(plotdata$ethnic) <- c("Other", "Non-Hispanic Black", "Non-Hispanic White", "Hispanic")
```

Here, I am looking at three psychiatric diagnoses and scoring them accordingly:

```{r}
cats = c("DX_01_Cat", "DX_02_Cat", "DX_03_Cat", "DX_04_Cat", "DX_05_Cat")
plotdata$anxiety <- apply(plotdata[,cats], 1, function(x)as.integer(any(grep("Anxiety Disorders",x))))
plotdata$depression <- apply(plotdata[,cats],1, function(x)as.integer(any(grep("Depressive Disorders",x))))
plotdata$bipolar <- apply(plotdata[,cats], 1, function(x)as.integer(any(grep("Bipolar and Related Disorders",x))))
```

Splitting the data by sex:
```{r}
girlsplotdata <- subset(plotdata, plotdata$Sex=="Females")
boysplotdata<- subset(plotdata, plotdata$Sex=="Males")
```

This is an example of one spline regression modelling of moderate-to-vigorous physical activity duration for males (based on checking whether linear fit or nonlinear fit was better for each model -- not shown here, available by request):

```{r}
nonlinearmvpamales <- gam(dur_day_total_MVPA ~ s(Age, bs="cr", k=15) + depression + PPS_M_04 + CIS_P_Scorev2 + wapct, data=boysplotdata, method="REML")
output <- summary(nonlinearmvpamales)
output
```

Creating a plot demonstrating males' MVPA, as an example: 

```{r}
aind <- seq(5,20,len=100)
df_pred_M <- expand.grid("Age"=aind, "PPS_M_04"=levels(boysplotdata$PPS_M_04))

##Since not all pubertal categories are observed in all age ranges, subset to the 5th and 95th percentile of age in each category for plotting 
df_age_pps_qs_M <- 
  boysplotdata %>% 
  group_by(PPS_M_04) %>% 
  summarize(q05 = quantile(Age, 0.05),
            q95 = quantile(Age, 0.95))

df_pred_M <- 
  left_join(df_pred_M, df_age_pps_qs_M, by="PPS_M_04") %>% 
  filter(Age >= q05, Age <= q95) %>% 
  dplyr::select(-q05,-q95) %>% 
  mutate(wapct = 0, 
         CIS_P_Scorev2 = 0, 
         ethnic = levels(boysplotdata$ethnic)[1],
         depression= 0)

##Predict average MVPA during the day at each pubertal score and add it to the dataframe we just created
fhat_mac_M <- predict(nonlinearmvpamales, newdata=df_pred_M, type='response',se.fit=TRUE)
##Add in point estimates
df_pred_M$yhat <- fhat_mac_M$fit
#Add in 95% CIs
df_pred_M$UB <- fhat_mac_M$fit + 1.96*fhat_mac_M$se.fit
df_pred_M$LB <- fhat_mac_M$fit - 1.96*fhat_mac_M$se.fit

##Make the plot (not including lower and upper CI bounds here for visual purposes)
plt_M <- 
  df_pred_M %>% 
  dplyr::select(Age, PPS_M_04, yhat, UB, LB) %>% 
  mutate(PPS_M_04 = factor(PPS_M_04, levels=2:5, labels=c("Level 1","Level 2","Level 3","Level 4"))) %>% 
  ggplot() + 
  geom_line(aes(x=Age,y=yhat, group=PPS_M_04, color=PPS_M_04)) +
  #geom_line(aes(x=Age,y=LB, group=PPS_M_04, color=PPS_M_04),lty=2) +
  #geom_line(aes(x=Age,y=UB, group=PPS_M_04, color=PPS_M_04),lty=2) +
  theme_classic() + ylab("[insert y-axis label]") + 
  ggtitle("(A) Boys [insert plot title]") +
  labs(color="Pubertal Status \n [insert labels]") +  
  theme(legend.position = c(0.8, 0.8)) + xlim(5,21)
```

```{r, echo=FALSE}
plt_M
```


Code to save my plot:

```{r, results=FALSE, message=FALSE}
ggsave(here::here(figure_path, "michelle_example_plot.jpeg"))
```
